---
title: "Bootstrapping ziCPUE data"
author: "Lucas A. Nell"
date: "March 31, 2016"
output:
  html_document:
    highlight: haddock
    number_sections: yes
    theme: default
    toc: yes
    toc_depth: 2
  pdf_document:
    highlight: haddock
    latex_engine: xelatex
mainfont: Helvetica
geometry: margin=1in
fontsize: 10pt
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, comment = '', prompt = FALSE)
knitr::read_chunk('bootstrapping.R')
```


# Initial setup

The following packages are required:

```{r packages, message = FALSE}
```


Read, clean, and add necessary columns to the table in `brett_example.xlsx`.

```{r inputData}
```


# Bootstrap resampling

For seamless bootstrapping with `dplyr`, the package `broom` provides the function
`boostrap`. However, this function samples from all rows regardless of grouping. I've
hacked the version of the function found [here](
https://cran.r-project.org/web/packages/broom/vignettes/bootstrapping.html) to sample 
within a given grouping as such: If, for example, you group by a variable called `type` 
and in the original sample, `type`s 1, 2, and 3  have sample sizes of 10, 20, and 30,
respectively. This function would randomly sample  exactly 10 items from the original 
dataset (with replacement) that are of `type` 1, 20 of `type` 2, and 30 of `type` 3.

```{r bootstrappingFuns}
```




Now for the actual bootstrapping. We'll do 1,000 simulations and calculate the 95% CI
via the "percentile" method.

```{r bootstrapping}
```



# Plotting


These are just some aesthetic pieces that I often employ. They're included here in case
others are interested.

```{r plotPreamble}
```


### Base plot

All following plots are based off this base one. Having no `geom`s, it just stores the
data and some aesthetics that we'll pass on to plots later.

```{r basePlot}
```

### Bootstrapping via `boot` in `ggplot2`

To have better control of bootstrapping within `ggplot2` plots, the following function
allows you to use the `boot` package for bootstrapping and CI calculations. The 
`ciMethod` parameter takes the following inputs, inherited from the `boot.ci` function:
"norm", "basic", "stud", "perc", or "bca". If parallel bootstrapping is desired, on 
Mac OSX, add `parallel = "multicore"` and on Windows add `parallel = "snow"` to the
`boot` function call; also add `ncpus = x`, where `x` is the number of available cores.

```{r ggBootCI_fun}
```


Now we'll employ the above function, using 1,000 simulations and calculating the 95% CI
via the "bca" method: bias-corrected and accelerated bootstrapping.

```{r bootPackagePlots}
```


### Plot on log~10~ scale

To create a plot on the log~10~ scale, we'll use the bootstrapping we did earlier via
the `stratBootstrap` function, located in the `CPUEbootSumm` data frame. Trying to 
transform the y-axis while calculating bootstrapped CI on the fly proved a fruitless
endeavor.

```{r WhitePlots}
```



