---
title: "Bootstrapping ziCPUE data"
author: "Lucas A. Nell"
date: "March 31, 2016"
output:
  html_document:
    highlight: haddock
    number_sections: yes
    theme: default
    toc: yes
    toc_depth: 2
  pdf_document:
    highlight: haddock
    latex_engine: xelatex
mainfont: Helvetica
geometry: margin=1in
fontsize: 10pt
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, comment = '', prompt = FALSE)
knitr::read_chunk('bootstrapping.R')
```


# Initial setup

The following packages are required:

```{r packages, message = FALSE}
```


Read, clean, and add necessary columns to the table in `brett_example.xlsx`.

```{r inputData}
```


# Bootstrap resampling

For seamless bootstrapping with `dplyr`, the package `broom` provides the function
`boostrap`.

I've grouped by `Year` to sample the same number of samples per year for each 
bootstrap resample as there were in the original sample.

If we represent each year's sampling as...
$$
n_1, n_2, n_3, \hspace{0.25em} ... \hspace{0.25em}, n_k
$$

where

$$
n_{total} = \sum_{i=1}^{k}{n_i}
$$

Then what we'll do below will maintain each of 
$n_1, \hspace{0.25em} ... \hspace{0.25em}, n_k$ 
(rather than just $n_{total}$) through all resamples.

We'll do 1,000 simulations and calculate the 95% CI via the "percentile" method.

```{r bootstrapping}
```



# Plotting


These are just some aesthetic pieces that I often employ. They're included here in case
others are interested.

```{r plotPreamble}
```


### Base plot

All following plots are based off this base one. Having no `geom`s, it just stores the
data and some aesthetics that we'll pass on to plots later.

```{r basePlot}
```

### Bootstrapping via `boot` in `ggplot2`

To have better control of bootstrapping within `ggplot2` plots, the following function
allows you to use the `boot` package for bootstrapping and CI calculations. The 
`ciMethod` parameter takes the following inputs, inherited from the `boot.ci` function:
"norm", "basic", "stud", "perc", or "bca". If parallel bootstrapping is desired, on 
Mac OSX, add `parallel = "multicore"` and on Windows add `parallel = "snow"` to the
`boot` function call; also add `ncpus = x`, where `x` is the number of available cores.

```{r ggBootCI_fun}
```


Now we'll employ the above function, using 1,000 simulations and calculating the 95% CI
via the "bca" method: bias-corrected and accelerated bootstrapping.

```{r bootPackagePlots}
```


### Plot on log~10~ scale

To create a plot on the log~10~ scale, we'll use the bootstrapping we did earlier via
the `stratBootstrap` function, located in the `CPUEbootSumm` data frame. Trying to 
transform the y-axis while calculating bootstrapped CI on the fly proved a fruitless
endeavor.

```{r WhitePlots}
```





# Session information


```{r session_info, include = TRUE, echo = TRUE, results = 'markup'}
devtools::session_info()
```



<!---
rmarkdown::render('bootstrapping.Rmd', output_file = 'bootstrapping.html')
-->