---
title: "Compiling data into a single data frame"
author: "Lucas A. Nell"
date: "May 22, 2016"
output:
  html_document:
    highlight: haddock
    number_sections: yes
    theme: default
    toc: yes
    toc_depth: 2
  pdf_document:
    highlight: haddock
    latex_engine: xelatex
mainfont: Helvetica
geometry: margin=1in
fontsize: 10pt
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, comment = '', prompt = FALSE)
knitr::read_chunk('compiling.R')
```

# Setup

## Required packages
```{r libraries, message = FALSE}
```

# Reading files

We are reading input files to `tbl_df`s named `Apal`, `Suwa`, `Pearl`, `Pasc`, and `Choc`.

```{r inputs}
```


# Standardizing data frames {.tabset}

Each data frame was unique in its column names and overall setup. Below are some goals
for the output data frames:

1. Contain as many of the following columns as possible: `Site`, `Date`, `vTagID`,
   `vSerial`, `PIT_Tag1`, `PIT_Tag2`, `TL_mm`, and `FL_mm` (equivalent columns  
   were converted using `rename` and selected for output using `select`) 
2. Convert the `FL` and `TL` columns from centimeters to millimeters
3. Standardize `Site` names and `Date` formats

## Apalach
*Note:* This one also had many unneeded columns, hence the `select` command at the 
beginning.
``` {r manApal}
```

## Suwannee
*Note:* This one also had some rows full of `NA`s, hence the `filter` command at the end.
``` {r manSuwa}
```

## Pearl
``` {r manPearl}
```

## Pascagoula
``` {r manPasc}
```


## Choctawhatchee

```{r manChoc}
```



# Gathering by PIT and combining

The below function uses `gather` to gather `PIT_Tag1` and `PIT_Tag2` into a single
`PIT_Tag` column, resulting in more rows but "tidier" data. This tidiness allows me to 
more easily combine the data frames into one and to match `PIT_Tag`s with `vTagID`s. 
The filter operation removes rows where `is.na(PIT_Tag)` and the associated
`vTagID`--`Date` combination has already been represented in another row. 
I removed `PIT_name` because it is useless for downstream processes.

``` {r gatherFun}
```


Now we run the function on all data frames and combine them into one. Since I made the 
data frames exactly the same above, combining them now is quite simple. I also
filtered rows that didn't have either a `PIT_Tag` or `vTagID`.
```{r runGather}
```






# Session information


```{r session_info, include = TRUE, echo = TRUE, results = 'markup'}
devtools::session_info()
```



<!---
rmarkdown::render('compiling.Rmd', output_file = 'compiling.html')
-->