---
title: "Compiling data into a single data frame"
author: "Lucas A. Nell"
date: "May 22, 2016"
output:
  html_document:
    highlight: haddock
    number_sections: yes
    theme: default
    toc: yes
    toc_depth: 2
  pdf_document:
    highlight: haddock
    latex_engine: xelatex
mainfont: Helvetica
geometry: margin=1in
fontsize: 10pt
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, comment = '', prompt = FALSE)
knitr::read_chunk('compiling.R')
```

# Setup

## Required packages
```{r libraries, message = FALSE}
```

# Reading files

We are reading input files to `tbl_df`s named `Apal`, `Suwa`, `Pearl`, `Pasc`, and `Choc`.

```{r inputs}
```


# Standardizing data frames {.tabset}

Each data frame was unique in its column names and overall setup. Below are some goals
for the output data frames:

1. Contain as many of the following columns as possible: `Site`, `Date`, `vTagID`,
   `vSerial`, `PIT_Tag1`, `PIT_Tag2`, `TL_mm`, and `FL_mm` (equivalent columns were 
   converted using `rename` and selected for output using `select`) 
2. Convert the `FL` and `TL` columns from centimeters to millimeters
3. Standardize `Site` names and `Date` formats

## Apalach
*Note:* This one also had many unneeded columns, hence the `select` command at the 
beginning.
``` {r manApal}
```

## Suwannee
*Note:* This one had no `vSerial` column, hence it's not present in the `select` command.
``` {r manSuwa}
```

## Pearl
*Note:* This one also had some rows full of `NA`s, hence the `filter` command at the 
beginning.
``` {r manPearl}
```

## Pascagoula
*Note:* This one only had one PIT tag column and none for `vSerial`, hence no `PIT_Tag2`
or `vSerial` in the `select` command.
``` {r manPasc}
```

## Choctawhatchee
*Note:* This one had no `vSerial` column, hence it's not present in the `select` command.
```{r manChoc}
```

## Yellow Escambia
```{r manYell}
```


## Panhandle rivers
*Note:* The `Internal` column is a logical vector specifying whether the tag for the 
captured individual was internal.
```{r manPanh}
```


## Pearl (ERDC)
*Note:* This one only had one PIT tag column, hence no `PIT_Tag2` in the `select` command.
```{r manErdc}
```




# Gathering by PIT and combining

## Gather function
The below function uses `gather` to gather `PIT_Tag1` and `PIT_Tag2` into a single
`PIT_Tag` column, resulting in more rows but "tidier" data. This tidiness allows me to 
more easily combine the data frames into one and to match `PIT_Tag`s with `vTagID`s. 
The filter operation removes rows where `is.na(PIT_Tag)` and the associated
`vTagID`--`Date` combination has already been represented in another row. 
I removed `PIT_name` because it is useless for downstream processes.
I lastly filtered rows that didn't have either a `PIT_Tag` or `vTagID` because they, too,
are useless.
``` {r gatherFun}
```


## *Side note*: `eval(as.name(x))`
The command `eval(as.name(x))` takes a character `x` and evaluates it as a named 
variable. Try the following code for an idea of what this does:
```{r evalCall, eval = FALSE}
duf <- data.frame(x = seq(10), y = seq(11, 20))
eval(as.name('duf'))

foo <- function(x){x ** 2}
eval(as.name('foo'))(5)
```


## Gathering all data frames and combining
Now we run the function on all data frames and combine them into one. 
Because I named all data frames with names ending in `'_df'`, I can filter for just
data frames in the `ls` command using the `pattern` argument.
On each of these character inputs, I use the `eval(as.name(x))` command discussed above.
Since I made the data frames nearly exactly the same above, combining them now is 
quite simple.
```{r runGather}
```


# Cleaning `Site` names
The original `Site` names were not standardized into "Systems" and "Subsystems". Using 
a data frame derived from the file `'site_name_corrections.csv'` in the `in_csv` folder
(provided to me by Bill Pine), I assigned standardized versions of each `Site` name.
I placed the `System` and `Subsystem` columns where `Site` was.
```{r cleanSites}
```




# Session information


```{r session_info, include = TRUE, echo = TRUE, results = 'markup'}
devtools::session_info()
```



<!---
rmarkdown::render('compiling.Rmd', output_file = 'compiling.html')
-->