---
title: "Compiling ziCPUE data"
author: "Lucas A. Nell"
date: "April 22, 2016"
output:
  html_document:
    highlight: haddock
    theme: journal
  pdf_document:
    highlight: haddock
    latex_engine: xelatex
mainfont: Helvetica
geometry: margin=1in
fontsize: 10pt
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, comment = '', prompt = FALSE)
```

# Setup

### Required packages:
```{r libraries, message = FALSE}
library("dplyr")
library("tidyr")
library("readxl")
library("stringr")
```

```{r dplyrSetup, echo = FALSE}
options(dplyr.print_max = 5, dplyr.width = Inf)
```

### Reading files
The `compiling_preamble.R` file inputs tables into `tbl_df`s named 
`Apal`, `Suwa`, `Pearl`, and `Pasc`. I did this separately to keep names private.
```{r inputs}
options(stringsAsFactors = FALSE)
source('compiling_preamble.R')
```


In `compiling_preamble.R`, files were read and nothing else. Examples for Excel and 
csv files, respectively:
```{r, eval = FALSE}
read_excel("FILENAME.xlsx")
read.csv("FILENAME.csv") %>% as.tbl
```



# Standardizing data frames

Each data frame was unique in its column names and overall setup. Below are some goals
for the output data frames:

1. Contain as many of the following columns as possible: `Site`, `Date`, `vTagID`,
   `vSerial`, `PIT_Tag1`, `PIT_Tag2`, `TL_mm`, and `FL_mm` (equivalent columns  
   were converted using `rename` and selected for output using `select`) 
2. Convert the `FL` and `TL` columns from cm to mm
3. Standardize `Site` names and `Date` formats
<!-- 4. Remove rows with missing `vTagID` -->

### Apalach
This one also had many unneeded columns, hence the `select` command at the 
beginning.
``` {r manApal}
Apal <- Apal %>%
  select(1:21) %>%
  rename(
    PIT_Tag1 = PIT_New, 
    PIT_Tag2 = Pit_Old,
    vTagID = V_TagID,
    vSerial = V_Serial) %>%
  mutate(
    FL_mm = FL_cm*10, 
    TL_mm = TL_cm*10,
    Date = as.Date(Date),
    Site = gsub('_', ' ', Site) %>% str_to_title
  ) %>%
  select(Site, Date, vTagID, vSerial, PIT_Tag1, PIT_Tag2, TL_mm, FL_mm)
```

### Suwannee
``` {r manSuwa}
Suwa <- Suwa %>%
  rename(
    vTagID = `ACOUSTIC TAG #`,
    PIT_Tag1 = `TAG 9    PIT TAG (ANTERIOR D FIN BASE)`,
    PIT_Tag2 = `TAG 10 EXTRA  OR AUX PIT TAG (INCL EXTRA PIT TAG 2)`,
    TL_mm = `TL mm`,
    FL_mm = `FL mm`,
    Site = `RIVER SYS`,
    Date = DATE
  ) %>%
  mutate(
    Date = as.Date(Date),
    Site = gsub('_', ' ', Site) %>% str_to_title
  ) %>%
  select(Site, Date, vTagID, PIT_Tag1, PIT_Tag2, TL_mm, FL_mm)
```

### Pearl
``` {r manPearl}
Pearl <- Pearl %>% 
  mutate(
    Date = as.Date(Date, format = '%m/%d/%Y'),
    Site="Pearl", 
    FL_mm=FL.cm*10, 
    TL_mm=TL.cm*10
  ) %>%
  rename(
    vTagID = Tel_tag_code, 
    PIT_Tag1 = Pit.Tag,
    PIT_Tag2 = old_Pit_tag
  ) %>%
  select(Site, Date, vTagID, PIT_Tag1, PIT_Tag2, TL_mm, FL_mm)
```

### Pascagoula
``` {r manPasc}
Pasc <- Pasc %>%
  mutate(
    Date = as.Date(Date_Tagged),
    Site = "Pascagoula",
    FL_mm = `FL(cm)`*10,
    TL_mm = `TL(cm)`*10
  ) %>%
  rename(
    vTagID = Tag_ID,
    PIT_Tag1 = `Pit Tag`
  ) %>%
  select(Site, Date, vTagID, PIT_Tag1, TL_mm, FL_mm)
```




# Gathering by PIT and combining

The below function uses `gather` to gather `PIT_Tag1` and `PIT_Tag2` into a single
`PIT_Tag` column, resulting in more rows but "tidier" data. This tidiness allows me to 
more easily combine the data frames into one and to figure out how many unique `PIT_Tag`
values there were for each `vTagID`. 
I removed `PIT_name` because it is useless for downstream processes.

``` {r gatherFun}
gatherByPIT <- function(df){
  df %>%
    gather(PIT_name, PIT_Tag, starts_with('PIT_Tag', ignore.case = FALSE)) %>%
    select(-PIT_name) %>%
    filter(!is.na(PIT_Tag))
}
```


Now we run the function on all data frames and combine them into one. Since I made the 
data frames exactly the same above, combining them now is quite simple.
```{r runGather}
allSites <- lapply(list(Apal, Suwa, Pearl, Pasc), gatherByPIT) %>% 
  bind_rows
allSites
```


I next want to figure out how many unique PIT Tag numbers are associated with 
`vTagID`s. The max appears to be 4.

```{r uniqueV_Tags_PITs}
allSites %>% 
  group_by(vTagID) %>%
  summarize(uniquePITs = length(unique(PIT_Tag))) %>%
  select(uniquePITs) %>% 
  table

allSites %>%
  filter(!is.na(vTagID)) %>%
  group_by(PIT_Tag) %>%
  summarize(uniquevTagIDs = length(unique(vTagID))) %>%
  select(uniquevTagIDs) %>% 
  table
```



# Matching `vTagID`s and `PIT_Tag`s

To easily access all identifying keys for each individual, I had to pair each `vTagID`
with all the `PIT_Tag`(s) associated with them. Because I had some instances of multiple 
`V_TagID`s for a given individual (based on `PIT_Tag`s), I also had to create a lookup
table for those instances and expand the `vTagID`--`PIT_Tag` table to account for these additional `PIT_Tag`s.


### `PIT_Tag`s paired with each `vTagID`



```{r vtPIT_lookup}
vTagID_PIT <- allSites %>%
  filter(!is.na(vTagID)) %>%
  distinct(vTagID, PIT_Tag) %>%
  select(vTagID, PIT_Tag)


lookupvTagID <- function(in_vTagID){
  vTagID_PIT %>%
    filter(vTagID == as.numeric(in_vTagID)) %>%
    select(PIT_Tag) %>%
    unlist %>% as.vector
}

```


### Equivalent `vTagID`s

Based on `PIT_Tag`s, some `vTagID` values represented the same individual. In case 
these need to be looked up, I've created a table to do just that. I first created a table
of these equivalent `vTagID`s, one row per equivalent pair/trio.
```{r equiv_vTagIDs_tmp}
equiv_vTagID_oneRow <- vTagID_PIT %>%
  group_by(PIT_Tag) %>%
  summarize(vTagIDs = paste(vTagID, collapse = ':')) %>%
  separate(vTagIDs, paste0('vTagID_', seq(3)), sep = ':', 
           fill = 'right') %>%
  filter(!is.na(vTagID_2)) %>%
  select(vTagID_1, vTagID_2, vTagID_3)
```

For easier lookup, I expanded the table so that every vTagID that was part of an 
equivalent pair/trio was present in the first column, `"prim"`.

```{r equiv_vTagID}
equiv_vTagID <- list(
  equiv_vTagID_oneRow %>% select(1, 2, 3) %>% 
    select(prim = 1, second = 2, tert = 3),
  equiv_vTagID_oneRow %>% select(2, 1, 3) %>% 
    select(prim = 1, second = 2, tert = 3),
  equiv_vTagID_oneRow %>% select(3, 1, 2) %>% 
    select(prim = 1, second = 2, tert = 3) %>% na.omit
  ) %>%
  bind_rows()
# `equiv_vTagID_oneRow` no longer needed...
rm(equiv_vTagID_oneRow)
```


















```{r valid_vTagIDs}
# A data frame of all observations with known vTagID or PIT_Tag
list((allSites %>% filter(!is.na(vTagID))), 
     (allSites %>%
        filter(is.na(vTagID)) %>% 
        filter(PIT_Tag %in% vTagID_PIT$PIT_Tag) %>%
        rowwise() %>%
        mutate(vTagID = vTagID_PIT$vTagID[vTagID_PIT$PIT_Tag == PIT_Tag]))) %>%
  bind_rows

```





``` {r, eval = F}

allSites %>%
  group_by(PIT_Tag) %>%
  summarize(vtag = length(na.omit(vTagID))) %>%
  select(vtag) %>%
  table


nrow(allSites); nrow(allSites %>% filter(!is.na(vTagID)))
sum(c(435, 66, 4))


```











