---
title: "Cross referencing"
author: "Lucas A. Nell"
date: "June 5, 2016"
output:
  html_document:
    highlight: haddock
    number_sections: yes
    theme: default
    toc: yes
    toc_depth: 2
  pdf_document:
    highlight: haddock
    latex_engine: xelatex
mainfont: Helvetica
geometry: margin=1in
fontsize: 10pt
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, comment = '', prompt = FALSE)
knitr::read_chunk('cross_referencing.R')
```


# Setup

## Required packages
```{r libraries, message = FALSE}
```

## Required objects

We need some of the objects created in the `identifiers` stage. For `masterCaps`, I 
explicitly specified the column types using `col_types = 'cDiiddc'` because by default 
it was making the `TL_mm` column an integer, rather than double like the original data
frame.

```{r read_ID_objects}
```


## Lookup functions {.tabset}

Below tabs have lookup functions for individual character inputs, and "vectorized" 
versions for vector input/output.

### Retrieve newest `vTagID`(s) from `vTagID`(s)

The function `new_vTagID` returns the newest `vTagID` from an input `vTagID` if that
`vTagID` is from an equivalent pair/trio. If it was a unique `vTagID`, the input is 
returned. `vec_new_vTagID` is the vectorized version that takes and returns a vector.

```{r newest_vTagID}
```

### Retrieve `vTagID`(s) from `PIT_Tag`(s)

> Below isn't really necessary because the only data frame with a `PIT_Tag` column doesn't 
> have any rows with a `PIT_Tag` but no `vTagID`.  I'm leaving it here for reference only.

The function `vTagID_from_PIT` returns the newest `vTagID` from an input `PIT_Tag` (the 
`PITs_w_vTag` data frame already has any `vTagID`s that are part of an equivalent 
pair/trio replaced with the newest one representing that individual).
`vec_vTagID_from_PIT` is the vectorized version that takes and returns a vector.

```{r vTagID_from_PIT}
```


# Reading, processing data for comparison to master


## `Eglin_List`...

I extracted the year from column `"Tag_description"` by splitting it by spaces, then 
retrieving the last string from that new character vector, then converting it to numeric.
I also removed the `old_V_TagID` column, since I'm only concerned with the most recent 
one.
I lastly made sure all `vTagID`s, if part of an equivalent pair/trio, were the newest one
representing that individual.

```{r read_Eglin}
```



## `All_GS_transmitters`...

This Excel sheet had many blank cells and cells with `"NA"`, each of which signifying
what in R would be `NA`. I had to change the blank ones to `"NA"` in Excel manually, 
since the `readxl::read_excel` function does not allow multiple representations of `NA`.
I also specified column types because some of the numbers were parsing incorrectly, 
throwing up errors.

I next removed the `"(F)"` from the `vTagID` column (note that inside the `gsub` function
you must double-escape parentheses using `"\\("` or `"\\)"`) and converted `vTagID` to
numeric.

The `River` column had extra spaces in some cells. Since none of the items in this 
column had >1 word, I simply removed all spaces. I then added `" River"` (`paste` 
adds a space between entries by default) and called it `Site`.

I lastly made sure all `vTagID`s, if part of an equivalent pair/trio, were the newest one
representing that individual.

This table did not have any `NA` `vTagID`s, but I would've included a check for that if 
it was necessary.

```{r read_allGS}
```



## `filtered_detections`

This differs from previous tables in that it's when individuals were detected by the
receiver array, not when they were captured.

The function `readr::read_rds` differs from other `readr` functions in that it returns
a simple data frame, rather than the "`tbl`" data frame, so I added a line for that.
I then assembled the `Date` column from the separate `Month`, `Day`, and `Year` columns.
Next I renamed the `Transmitter` column to `vTagID`, removed unnecessary columns, 
replaced dropped `vTagID`s, and arranged row by `River`, `Date`, `Receiver`, and 
`vTagID`.

```{r read_filtered_detections}
```




# Comparing `Eglin` to master

## Group by year

Since I only have `Year` (rather than full dates) from the `Eglin` data frame, I'll group 
by `Year` and compute how many times that individual was caught in that year. 
I then created the `Year_vTagID` column, which contains the `Year` and `vTagID` pasted
together and separated by an underscore. By comparing this column across data frames, I
can assess what `Year`–`vTagID` combinations we're missing from `masterCaps`.

```{r comp_Eglin_makeDFs}
```


## Shared rows

By using the `Year_vTagID` column, I find that they share only one `vTagID`–`Year`
combination.

```{r comp_Eglin_sharedRow}
```



## `vTagID`s not in master

Now I check how many unique `vTagID`s present in `Eglin` are not present in the master
capture data frame. We find that 
`r length(unique(Eglin$vTagID)[! unique(Eglin$vTagID) %in% masterCaps$vTagID])` out of 
`r length(unique(Eglin$vTagID))` unique `vTagID`s in `Eglin` are not present in the 
master capture data frame.

```{r comp_Eglin_vTagIDs_notInMaster}
```





# Comparing `allGS` to master

## Make `Date`–`vTagID` column

Similar to what I did with the `Eglin` data frame, I created a `"Date_vTagID"` column
that pastes together the `Date` and `vTagID` columns.

```{r comp_allGS_makeDFs}
```


## Rows not in master

Using the `Date_vTagID` column, I find that a little over half of the 
`r nrow(allGS)` rows in `allGS` are not present in
the master capture data frame. 

```{r comp_allGS_rows_notInMaster}
```





# Comparing `filtDet` to master

This data frame contains dates where individuals were detected on a receiver 
array, *not* when they were captured, as in the master capture data frame (`masterCaps`).
Thus I will not compare `Date`–`vTagID` combinations, just `vTagID`s.


## vTagIDs from `filtDet` in `masterCaps`

First, I find that 
`r length(unique(filtDet$vTagID)[! unique(filtDet$vTagID) %in% masterCaps$vTagID])`
out of 
`r length(unique(filtDet$vTagID))` unique `vTagID`s from `filtDet` are not already in the 
master capture data frame.

```{r comp_filtDet_in_masterCaps}
```


## vTagIDs from `masterCaps` in `filtDet`

I also find that 
`r length(unique(masterCaps$vTagID)[ unique(masterCaps$vTagID) %in% filtDet$vTagID])`
out of 
`r length(unique(masterCaps$vTagID))`
unique `vTagID`s from `masterCaps` that are already in `filtDet`.

```{r comp_masterCaps_in_filtDet}
```


## `filtDet` to `masterCaps` comparison table

Table of unique `vTagID`s from `filtDet` and whether they're present in `masterCaps`. If
an individual was caught in multiple rivers, I pasted them together, separated by `':'`.

```{r comp_filtDet_in_masterCaps_Table}
```


## `masterCaps` to `filtDet` comparison table

For this table, I simply recreated `masterCaps` with an extra column indicating whether 
the vTagID was present in the `filtDet` data frame.

```{r comp_masterCaps_in_filtDet_Table}
```


## Saving

If you want to save these comparison tables as csv files:
```{r writeTabs, eval = FALSE}
write_csv(filtDet_inMaster, './csv_out/filtDet_inMaster.csv')
write_csv(masterCaps_inFD, './csv_out/masterCaps_inFD.csv')
```





# Session information


```{r session_info, include = TRUE, echo = FALSE, results = 'markup'}
devtools::session_info()
```





<!---
rmarkdown::render('cross_referencing.Rmd', output_file = 'cross_referencing.html')
-->