---
title: "Cross referencing"
author: "Lucas A. Nell"
date: "June 5, 2016"
output:
  html_document:
    highlight: haddock
    number_sections: yes
    theme: default
    toc: yes
    toc_depth: 2
  pdf_document:
    highlight: haddock
    latex_engine: xelatex
mainfont: Helvetica
geometry: margin=1in
fontsize: 10pt
runtime: shiny
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, comment = '', prompt = FALSE)
knitr::read_chunk('cross_referencing.R')
```


# Setup

## Required packages
```{r libraries, message = FALSE}
```

## Required objects

We need the objects created in the `identifiers` stage.

```{r read_ID_objects}
```


## Lookup functions {.tabset}

Below tabs have lookup functions for individual character inputs, and "vectorized" 
versions for vector input/output.

### Retrieve newest `vTagID` from a vector of `vTagID`s

```{r newest_vTagID}
```

### Retrieve `vTagID` from `PIT_Tag`

Below isn't really necessary because the only data frame with a `PIT_Tag` column doesn't 
have any rows with a `PIT_Tag` but no `vTagID`. 
I'm leaving it here for reference only.
```{r vTagID_from_PIT}
```


# Reading, processing data for comparison to master


## `Eglin_List`...

```{r read_Eglin}
```


## `All_GS_transmitters`...

```{r read_allGS}
```

## `filtered_detections`

This differs from previous tables in that it's when individuals were detected by the
receiver array, not when they were captured.
```{r read_filtered_detections}
```




# Comparing `Eglin` to master



# Comparing `allGS` to master



# Comparing `filtDet` to master














# Session information


```{r session_info, include = TRUE, echo = TRUE, results = 'markup'}
devtools::session_info()
```





<!---
rmarkdown::render('cross_referencing.Rmd', output_file = 'cross_referencing.html')
-->