---
title: "Identifiers: Assembling master list and dropped `vTagID`s"
author: "Lucas A. Nell"
date: "May 22, 2016"
output:
  html_document:
    highlight: haddock
    number_sections: yes
    theme: default
    toc: yes
    toc_depth: 2
  pdf_document:
    highlight: haddock
    latex_engine: xelatex
mainfont: Helvetica
geometry: margin=1in
fontsize: 10pt
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, comment = '', prompt = FALSE)
knitr::read_chunk('identifiers.R')
```

# Setup

## Required packages
```{r libraries, message = FALSE}
```

## Required objects

We need the `cleanSites` data frame created in the compiling stage. The `rm` command 
removes all other objects created in `compiling.R`.
```{r sourceCompiling}
```



# Matching `vTagID`s and `PIT_Tag`s

To easily access all identifying keys for each individual, I had to pair each `vTagID`
with all the `PIT_Tag`(s) associated with them, and vice versa.

## Counting unique matches {.tabset}

I first want to figure out how many unique, non-`NA` `PIT_Tag`s are associated with 
`vTagID`s, and vice versa. 

### `PIT_Tag`s per `vTagID`
Table of the number of unique, non-`NA` `PIT_Tag`s per `vTagID`
```{r unique_PITs_per_vTag}
```

### `vTagID`s per `PIT_Tag`
Table of the number of unique, non-`NA` `vTagID`s per `PIT_Tag`
```{r unique_vTags_per_PIT}
```


## Maximum matches

I'll create an object to store the maximum value for both tables above for later.

```{r maxUniques}
```




## `PIT_Tag`--`vTagID` pairs

I needed to create a data frame of all unique `vTagID`--`PIT_Tag` pairs, so that this 
data frame doesn't have to be created for every single lookup (which would be a real
time-waster).
Because some `PIT_Tag`s match with multiple `vTagID`s, I will expand this data frame
later to account for this.

```{r vTag_PIT_lookup_DF}
```


## Lookup functions

I next created functions to look up all equivalent `PIT_Tags` for a given `vTagID`, 
and vice versa, from the `vTagID_PIT` data frame. 
I will later be adding rows to the `vTagID_PIT` reference data frame, and even 
though I am creating these functions now, they will use the updated version of
`vTagID_PIT` if they're called after the changes are made to the data frame.
Each of these functions uses an inner function that works for a single input
`vTagID`/`PIT_Tag`, then uses the `sapply` function to apply that inner function to
a vector.

```{r vTag_PIT_lookup_Funs}
```





# Equivalent `vTagID`s

Based on `PIT_Tag`s, some individuals were represented by multiple `vTagID` values. These
were individuals where the tag fell off/out and were replaced with a new one.

To simplify matters, I'll make a lookup data frame, and in the final output data frame,
I'll replace any `vTagID`s that are part of an equivalent pair/trio with the most recent
`vTagID` representing that individual.


## Newest `vTagID` lookup

The below function will return the most recent `vTagID` from an input vector of `vTagID`s.
*Note*: This function does not retrieve the newest `vTagID` for a given *individual*,
just from the provided `vTagID`s.

``` {r newest_vTagID_fun}
```


## Equivalent `vTagID` lookup data frame

This lookup data frame can be used to find the newest equivalent `vTagID` for a given
input `vTagID`. I first created a data frame of these equivalent `vTagID`s, one row 
per equivalent pair/trio, including the `PIT_Tag` that allowed me to match them.

```{r equiv_vTagID_PIT}
```

For easier lookup of whether a given `vTagID` is part of an equivalent pair/trio, and, if
so, what the newest `vTagID` is for that individual, the final lookup data frame is a
"longer" version of the one above.
In this data frame, every `vTagID` that is part of an equivalent pair/trio 
is present in the first column, `"input"`. The newest `vTagID` for that individual 
is in the `"newest"` column.

The code chunk below does the following:

1. Create a data frame using `data_frame` from package `tibble`, which is faster and
   automatically adds `tbl_df` class to output.
2. Add the following named columns to the data frame:
    a. `"input"`: The first 3 columns of `equiv_vTagID_PIT` turned into a single vector.
    b. `"newest"`: `apply` the `findNewest_vTagID` function across rows of the
       `equiv_vTagID_PIT` data frame, but only including the first 3 columns (the 
       ones with `vTagID`s). This returns a vector of the newest `vTagID`s per pair/trio.
       I also used `repeat` on `newest_vTagIDs` with the option `each = 3`, to make 
       this vector the same length as that in the `"input"` entry and to make 
       them refer to the same individual at all locations across both vectors.
3. Remove any rows where `input` is not present.

```{r equiv_vTagID}
```




## Retrieving full information for captures including equivalent `vTagID`s

To check on equivalent `vTagID`s, below is how to create a csv file
of each capture containing an equivalent `vTagID`.

```{r equivRows, eval = FALSE}
equivRows <-
    cleanSites %>%
    filter(vTagID %in% equiv_vTagID$input, 
           PIT_Tag %in% equiv_vTagID_PIT$PIT_Tag) %>% 
    arrange(System, Subsystem, PIT_Tag, vTagID)

write_csv(equivRows, './csv_out/equivalent_vTagIDs.csv')
```



## Adjusting `vTagID`--`PIT_Tag` lookup data frame for dropped vTags

I need to next adjust the `vTagID`--`PIT_Tag` lookup data frame to account for the fact 
that some individuals are represented by multiple `vTagID`s because of dropped vTags that
were replaced. 
For each of these individuals, I'm adding rows containing...

1. The newest `vTagID` associated with them (column `"vTagID"`)
2. All `PIT_Tag`s associated with any `vTagID` that has ever been on that individual
    (column `"PIT_Tag"`)

```{r expand_vTagID_PIT_NewRows}
```


Now I add these rows to the previous `vTagID_PIT` data frame, filtering for unique
pairings of `vTagID` and `PIT_Tag`, and arranging by `vTagID` in ascending order.

```{r expand_vTagID_PIT}
```




# Final data frame of all `PIT_Tag`s with a `vTag`

I first assigned an object for the maximum number of `vTagID`s for any given `PIT_Tag`.

```{r max_vTagIDs}
```

I next made the data frame of `PIT_Tag`s and `vTagID`s into a data frame of all `PIT_Tag`s
with a `vTagID`. If more than one `vTagID` is associated with a given `PIT_Tag`, I used 
the newest `vTagID`. I lastly added a `"River"` column by referencing the `cleanSites`
data frame.

```{r PITs_w_vTag}
```

Now I can save this as a csv file.

```{r save_PITs_w_vTag, eval = FALSE}
write_csv(PITs_w_vTag, './csv_out/PITs_w_vTag.csv')
```



# Final data frame of dropped `vTagID`s

To change the `equiv_vTagID` lookup data frame to a data frame of just dropped 
`vTagID`s, I removed rows where column `"input"` contains a `vTagID` located anywhere in
the `"newest"` column. I then renamed the columns for clarity in this context.

```{r dropped_vTagIDs}
```

And to save it:

```{r save_dropped_vTagIDs, eval = FALSE}
write_csv(dropped_vTagIDs, './csv_out/dropped_vTagIDs.csv')
```



# Master data frame of captures

## Making master data frame only use newest `vTagID`s

I now replace all instances in the master observation data frame of dropped `vTagID`s 
with the newest `vTagID` that represents the same individual.

```{r newestID_master}
```


## Making master data frame contain only valid `vTagID`s

By cross-referencing with the data frame of known `vTagID`s for `PIT_Tag`s, I filled
in some rows that don't have `vTagID`s, but do have `PIT_Tag`s. I combined these rows 
with those from `cleanSites` that already have valid (i.e., non-`NA`) `vTagID`s to create
a data frame of all observations with known identifiers.

To avoid creating temporary objects, both sets of rows were created on the fly as data 
frames within a list.

```{r knownID_master}
```


Now I can save this as a csv file.

```{r save_masterCaps, eval = FALSE}
write_csv(masterCaps, './csv_out/masterCaps.csv')
```





# Session information


```{r session_info, include = TRUE, echo = FALSE, results = 'markup'}
devtools::session_info()
```



<!---
rmarkdown::render('identifiers.Rmd', output_file = 'identifiers.html')
-->